---
title: "Project 1 Reproducible Research"
output:
  html_document: default
  pdf_document: default
---
```{r}
library(ggplot2)
```
##Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")

```

##What is mean total number of steps taken per day?


```{r}

stepsday <- tapply(activity$steps,activity$date,sum,na.rm = TRUE)
```

####Make a histogram of the total number of steps taken each day
```{r}
qplot(stepsday,xlab= 'Total Steps',ylab= 'Number of days',binwidth = 500)
```

####Calculate and report the mean and median total number of steps taken per day
```{r}
meanbyday <- mean(stepsday)
medianbyday <- median(stepsday)
```
* Mean:  `r meanbyday`

* Median:`r medianbyday`


##What is the average daily activity pattern?

####Make a time series plot
```{r}
stepsday <- aggregate(x=list(meansteps=activity$steps),by=list(interval = activity$interval),mean, na.rm=TRUE )
```

####Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
maxvalue<- stepsday[which.max(stepsday$meansteps),]
```

* Interval and max steps: `r maxvalue`

#Imputing missing values

####Calculate and report the total number of missing values in the dataset

```{r}
missing <- length(which(is.na(activity$steps)))

```
 
 * Missing values: `r missing`
 
####Devise a strategy for filling in all of the missing values in the dataset.

```{r}
activityimp <- merge(activity,stepsday,by= "interval")
activityimp <- within(activityimp, steps<- ifelse(is.na(activityimp$steps),activityimp$meansteps,activityimp$steps))
```

####Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
totstepsday <- tapply(activityimp$steps,activityimp$date,sum,na.rm = TRUE)
```

####Make a histogram of the total number of steps taken each day

```{r}
qplot(totstepsday,xlab= 'Total Steps',ylab= 'Number of days',binwidth = 500)
```

####Calculate and report the mean and median total number of steps taken per day

```{r}
totmeanbyday <- mean(totstepsday)
totmedianbyday <- median(totstepsday)
```

* Mean:  `r totmeanbyday`
* Median:`r totmedianbyday`

##Are there differences in activity patterns between weekdays and weekends?

####Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
activityimp$dateType <-  ifelse(as.POSIXlt(activityimp$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

####Make a panel plot containing a time series plot

```{r}
averagedActivityimp <- aggregate(steps ~ interval + dateType, data=activityimp, mean)
ggplot(averagedActivityimp, aes(interval, steps))+  
        geom_line() + 
        facet_grid(dateType ~ .)+ 
        xlab("5-minute interval")+  
        ylab("avarage number of steps")
